The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2010-03-04 script/en/chapters/numbertheory.tex
Some more minor corrections in the script (German and English); Problem building English script.

From algebra theorems it follows that essential parts of the conventional
calculation rules are kept when we proceed to modular calculations over
a basic set $\mathbb{Z}$.  For example, addition remains commutative. 
The same goes for multiplication modulo $m$. The result of a
division\footnote{%
When dividing modulo $m$\index{Division modulo $n$} we can one use
co-prime\index{Number!co-prime} numbers because because other numbers
have the same property as zero. See footnote
\ref{ftn-mod6} in chapter~\ref{addmult}.
\label{ftn-res-divmodn}
\label{ftn-zth-divmodn}
} is not a fraction but rather a whole number between $0$ and $m-1$.

@@

\newpage
\begin{center}
\fbox{\parbox{15cm}{
    \emph{Seneca\index{Seneca}\footnotemark:}\\
    The way of theory is long, it is short and effective by examples.
}}
\end{center}
\addtocounter{footnote}{0}
\footnotetext{%
     Lucius Annaeus Seneca, philosophical writer and poet, 4 B.\:C. $-$ 65 A.\:D.}

@@

Within the article \cite{nt:Bourseau2002} Dirk Fox\footnote{%
His company Secorvo Ltd also delivered a statement on the recommendation for key
length selection published by the GISA\index{GISA} (German Information
Security Agency). Chapter 2.3.1 of this statement contains a competent
and understandable discussion of RSA security (this document
exists -- to my knowledge -- only in German):\\
\href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
{\texttt{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}}
}
published his prognosis of an almost linear factorization progression,
if all influencing factors are included: Each year the module length feasible for
factorization increases by 20 bit on average. So his forecast was below the
more optimistic estimations of GISA and NIST.

@@

\subsection{Status regarding factorization of concrete large numbers}
\label{NoteFactorization}

@@

\noindent Below the last records are explained in more detail.
The two methods, GNFS and SNFS, used to do so are shortly illustrated at the 
following web pages:
\index{General Number Field Sieve (GNFS)}
\index{Special Number Field Sieve (SNFS)}
\vspace{-10pt}
\begin{itemize}
\item[]
      \href{http://en.wikipedia.org/wiki/Special_number_field_sieve}
   {\texttt{http://en.wikipedia.org/wiki/Special\_number\_field\_sieve}} \\
      \href{http://en.wikipedia.org/wiki/General_number_field_sieve}
   {\texttt{http://en.wikipedia.org/wiki/General\_number\_field\_sieve}}
\end{itemize}
%\vspace{-10pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
\href{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}
{\texttt{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}}}
factorized a 158-digit decimal number into its both prime factors (these are
build with 73 and 86 decimal digits) using the GNFS method (General Number
Field Sieve)\index{General Number Field Sieve (GNFS)}.

@@

The task of the researchers from Bonn was not initiated by a challenge, but
they wanted to find the last prime factors of the integer $2^{953} - 1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
                 {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}}).

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g.\ using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorization of a Number}. \\
CrypTool can factorize in a reasonable time numbers no longer than 250 bit.
Numbers bigger than 1024 bits are currently not accepted by CrypTool.
}.
The next three prime factors were found by P.~Zimmerman\footnote{%
\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}}, 
T.~Grandlund\footnote{\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/}}
and R. Harley during the years 1999 and 2000 using the elliptic curve factorization method.

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \href{http://www.loria.fr/~{}zimmerma/records/gnfs158}
       {\texttt{http://www.loria.fr/~{}zimmerma/records/gnfs158}} \\
          \href{http://www.crypto-world.com/FactorRecords.html}
       {\texttt{http://www.crypto-world.com/FactorRecords.html}} \\
          \href{http://www.crypto-world.com/announcements/c158.txt}
       {\texttt{http://www.crypto-world.com/announcements/c158.txt}}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
          \href{http://www.loria.fr/~zimmerma/records/rsa160}
       {\texttt{http://www.loria.fr/\~{}zimmerma/records/rsa160}} \\
          \href{http://www.loria.fr/~zimmerma/records/factor.html}
       {\texttt{http://www.loria.fr/\~{}zimmerma/records/factor.html}} \\
          \href{http://www.crypto-world.com/FactorWorld.html}
       {\texttt{http://www.crypto-world.com/FactorWorld.html}}
} 
factorized a 160-digit number into its both prime factors (these are build
with each 80 decimal digits) using the GNFS method (General Number Field
Sieve)\index{General Number Field Sieve (GNFS)}.

The  computations for the factorization of RSA-160 also took place at the 
German Information Security Agency (GISA) in Bonn\footnote{%
Every year the GISA\index{GISA} creates a paper to describe which 
crypto algorithms are feasible to generate digital signatures according
to the German signature law -- under participation of experts from
economy and science. To review signature methods based on the
factorization problem the GISA also co-operates with researchers from the
University of Bonn.
Further information about crypto algorithms can be found on the web page of GISA:
   \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
{\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}}.
}.

@@

The task of the researchers was not initiated by a challenge, but
they wanted to find the last prime factors of the Mersenne number $2^{1039}+1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
                 {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}\\
Cunningham table: \href{http://homes.cerias.purdue.edu/~ssw/cun/pmain1206}
                {\texttt{http://homes.cerias.purdue.edu/\~{}ssw/cun/pmain1206}}\\
The numbers in the Cunningham table have the following syntax:\\
``(2,n)-'' means $2^{n}-1$;~~~
``(2,n)+'' means $2^{n}+1$.\\
To describe the magnitude one writes $p<n>$ or $c<n>$: 
``n'' is the number of decimal digits and ``p'' and ``c'' tell,
whether the number is prime or composite.\\
$2^{1039}-1 = p7 * c307 = p7 * p80 * p227$ \\
It is explained more precisely at the CUN page:\\
``2651+ means $2^{651} + 1$ and the size (c209 means 209 decimal digits)
of the
number which was factored.  Then come the new factor(s), the discoverer and
the method used.  Recently, only the multiple polynomial quadratic sieve
(ppmpqs), the elliptic curve method (ecm) and the number field sieve (nfs)
have been used.  `hmpqs' stands for hypercube multiple polynomial quadratic
sieve.  Under `new factors', `p90' means a 90-digit prime and `c201' is a
201-digit composite number.''.
}).\\

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \href{http://www.loria.fr/~zimmerma/records/21039-}
       {\texttt{http://www.loria.fr/\~{}zimmerma/records/21039-}} \\
          \href{http://www.crypto-world.com/announcements/m1039.txt}
       {\texttt{http://www.crypto-world.com/announcements/m1039.txt}} \\
          \href{http://www.crypto-world.com/FactorAnnouncements.html}
       {\texttt{http://www.crypto-world.com/FactorAnnouncements.html}} \\
          \href{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?detailjahr=2007&detail=160}
       {\texttt{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?}} \\
          \href{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?detailjahr=2007&detail=160}
       {\texttt{\quad \quad \quad detailjahr=2007\&detail=160}}\end{itemize}
\end{minipage}

@@

In his paper ``Circuits for integer factorization: a proposal'' 
(\href{http://cr.yp.to/djb.html}{\texttt{http://cr.yp.to/djb.html}}),
published November 2001, D.~J.\ Bernstein \cite{nt:Bernstein2001} 
addresses the problem of
factorizing large integers. Therefore, his results are of relevance from a
RSA point of view.  As a main result Bernstein claims that the
implementation of the General Number Field Sieve algorithm (GNFS)
 \index{General Number Field Sieve (GNFS)} can be improved to factor, with
the same effort as before, integers with three times more digits.

@@

\href{http://www.rsasecurity.com/}{RSA Security's} analysis of the Bernstein
paper \cite{nt:RSA Security 2002} from April, 8 2002 also -- as expected --
concludes, that RSA is still not compromised.

@@

\vskip +12pt
\noindent\begin{minipage}{\textwidth}
Links:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://cr.yp.to/djb.html}
               {\texttt{http://cr.yp.to/djb.html}} \\
          \href{http://www.counterpane.com/crypto-gram-0203.html\#6}
               {\texttt{http://www.counterpane.com/crypto-gram-0203.html\#6}} \\
          \href{http://www.math.uic.edu}
               {\texttt{http://www.math.uic.edu}}
\end{itemize}
\end{minipage}

@@

\noindent Current research results on this topic can be found at:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.mersenne.org/}
               {\texttt{http://www.mersenne.org/}} \\
          \href{http://fatphil.org/maths/AKS/}
               {\texttt{http://fatphil.org/maths/AKS/}} Original paper in English\\
          \href{http://ls2-www.cs.uni-dortmund.de/lehre/winter200203/kt/material/primes.ps}
               {\texttt{http://ls2-www.cs.uni-dortmund.de/lehre/winter200203/kt/material/primes.ps}}\\
	       \hspace*{2em}Good explanation in German by Thomas Hofmeister. 
\end{itemize}
\vskip +10 pt

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{Chapter_ElementaryNT_12}{}
\section[The RSA procedure with actual numbers]{The RSA procedure with actual numbers\footnotemark}
\footnotetext{%
    \index{Sage}%
    \index{Nguyen, Minh Van }%
    Additional material: Minh Van Nguyen, ``Number Theory and the RSA
    Public Key Cryptosystem'', 2009. An introductory tutorial on using
    Sage to study elementary number theory and public key
    cryptography. A didactically very clear article about some basic
    number theory and Sage usage. 
    \href{http://nguyenminh2.googlepages.com/numtheory-crypto-sage.pdf}
    {\url{http://nguyenminh2.googlepages.com/numtheory-crypto-sage.pdf}}.
}
\label{rsaconcrete}\index{RSA}
Having described above \hyperlink{RSA}{how the RSA procedure works}, we will now work through the steps using actual, but small, numbers.

@@

The task is taken from \cite[Exercise 4.6]{nt:Stinson1995}\index{Stinson 1995}:
The pure solution has been published by Prof. Stinson at 
\href{http://www.cacr.math.uwaterloo.ca/~dstinson/solns.html}{\texttt{http://www.cacr.math.uwaterloo.ca/\~{}dstinson/solns.html}}.\footnote{or
\href{http://bibd.unl/~stinson/solns.html}{\texttt{http://bibd.unl/\~{}stinson/solns.html}}.}

@@

The file ``primroots.dat'' is a database of all primitive roots of all primes
between 1 and 100,000 inclusive. It is a very large file (about 1 GB
uncompressed, and 285 MB compressed with bzip2). You can find the file
at
   \href{http://sage.math.washington.edu/home/mvngu/doc/primitive-roots/primroots.dat.bz2}
   {\url{http://sage.math.washington.edu/home/mvngu/doc/primitive-roots/primroots.dat.bz2}}.

@@

\begin{enumerate}
   \item \hypertarget{knott}{}  \index{Knott, Ron}
         Ron Knott's Fibonacci \index{Fibonacci} page, \\
         Here, everything revolves around Fibonacci numbers.\\
         \href{http://www.mcs.surrey.ac.uk/personal/R.Knott/Fibonacci/fib.html}
         {\texttt{http://www.mcs.surrey.ac.uk/personal/R.Knott/Fibonacci/fib.html}}

   \item CrypTool, \index{CrypTool} \\
         Open source e-learning software to illustrate cryptography and cryptanalysis\\
         \href{http://www.cryptool.de}{\texttt{http://www.cryptool.de}},\\
         \href{http://www.cryptool.org}{\texttt{http://www.cryptool.org}},\\ 
         \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}

   \item Mathematica, \index{Mathematica} \\
         Commercial mathematics package \\
         \href{http://www.wolfram.com}{\texttt{http://www.wolfram.com }}

   \item LiDIA, \index{LiDIA} \\
         Extensive library containing number-theory functions and the LC interpreter \\
         \href{http://www.informatik.tu-darmstadt.de/TI/LiDIA}
         {\texttt{http://www.informatik.tu-darmstadt.de/TI/LiDIA}}

   \item BC, \index{BC} \\
         Interpreter with number-theory functions \\
         \href{http://www.gnu.org/software/bc/bc.html}
         {\texttt{http://www.gnu.org/software/bc/bc.html}}

   \item Pari-GP, \index{Pari-GP} \\
         Excellent, fast, free interpreter with number theoretical functions.\\
         \href{http://pari.math.u-bordeaux.fr/}
              {\texttt{http://pari.math.u-bordeaux.fr/}}\\
         \href{http://en.wikipedia.org/wiki/PARI/GP}
              {\texttt{http://en.wikipedia.org/wiki/PARI/GP}}\\
         Resources for PARI/GP at Karim Belabas's website:\\
         \href{http://www.math.u-bordeaux.fr/~belabas/pari/}
              {\texttt{http://www.math.u-bordeaux.fr/\~{}belabas/pari/}}

   \item \index{Munchenbach@M\"unchenbach, Carsten}
         Only after I had completed this article, did I come across the
         website of Mr.\ M\"unchenbach, which interactively and didactically
         uses elementary number theory to provide a sophisticated
         description of the fundamental mathematical thought processes. 
         It was created for a teaching project in the 11th grade 
         of the technical grammar school (unfortunately only available in German): \\
         \href{http://www.hydrargyrum.de/kryptographie}
         {\texttt{http://www.hydrargyrum.de/kryptographie}}

   \item Web site of Mr.~Wagner, who is responsible for the development
	 of the curriculum of computer science in one of the German federal 
         states (L\"ander). Here you can get hold of a collection of texts
	 and (Java-)\discretionary{}{}{}programs (available only in German):\\
         \href{http://www.saar.de/~awa/kryptolo.htm}
         {\texttt{http://www.saar.de/\~{}awa/kryptolo.htm}}

   \item GISA, \index{GISA} \\
         German Information Security Agency \\
         \href{http://www.bsi.bund.de}{\texttt{http://www.bsi.bund.de}}

   \item Factorization records and challenges\index{Factorization!factoring records},\\
         \href{http://www.crypto-world.com/}
           {\texttt{http://www.crypto-world.com/}} \\
         \href{http://www.crypto-world.com/FactorWorld.html}
           {\texttt{http://www.crypto-world.com/FactorWorld.html}}, page by Scott Contini \\
         \href{http://www.loria.fr/~zimmerma/records/factor.html}
           {\texttt{http://www.loria.fr/\~{}zimmerma/records/factor.html}} \\
         \href{http://www.tutorgig.com/ed/RSA\_number}
           {\texttt{http://www.tutorgig.com/ed/RSA\_number}} \\

         \href{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}
           {\texttt{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}} \\
         \href{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}
           {\texttt{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html, 2002-01}} \\
         \href{http://www.loria.fr/~zimmerma/records/rsa160}
           {\texttt{http://www.loria.fr/\~{}zimmerma/records/rsa160}} \\

         \href{http://www.rsa.com/rsalabs/node.asp?id=2092}
           {\texttt{http://www.rsa.com/rsalabs/node.asp?id=2092}}\\

   \item The Cunningham Project, \index{Cunningham project}\\ 
         \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
         {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}

   \item Sage, \index{Sage} \\
         Excellent, open source computer algebra system with Python as script language,
	 used to build the code samples in this chapter.
	 See the introduction in chapter~\ref{s:appendix-using-sage}.\\
         \href{http://www.sagemath.org/}
              {\texttt{http://www.sagemath.org/}}\\
         \href{http://en.wikipedia.org/wiki/Sage_%28mathematics_software%29}
              {\texttt{http://en.wikipedia.org/wiki/Sage\_\%28mathematics\_software\%29}}\\

>>>>

From algebra theorems it follows that essential parts of the conventional
calculation rules are kept when we proceed to modular calculations over
a basic set $\mathbb{Z}$.  For example, addition remains commutative. 
The same goes for multiplication modulo $m$. The result of a
division\footnote{%
The division modulo $m$\index{Division modulo $n$} is only defined for
numbers co-prime\index{Number!co-prime} to $m$ because other numbers
have the same property as zero. See footnote
\ref{ftn-mod6} in chapter~\ref{addmult}.
\label{ftn-res-divmodn}
\label{ftn-zth-divmodn}
} is not a fraction but rather a whole number between $0$ and $m-1$.

@@

\newpage
\begin{center}
\fbox{\parbox{15cm}{
    \emph{Seneca\index{Seneca}\footnotemark:}\\
    The way of theory is long  ---  it is short and effective by examples.
}}
\end{center}
\addtocounter{footnote}{0}
\footnotetext{%
     Lucius Annaeus Seneca, philosophical writer and poet, 4 B.\:C. $-$ 65 A.\:D.}

@@

Within the article \cite{nt:Bourseau2002} Dirk Fox\footnote{%
His company Secorvo Ltd also delivered a statement on the recommendation for key
length selection published by the GISA\index{GISA} (German Information
Security Agency). Chapter 2.3.1 of this statement contains a competent
and understandable discussion of RSA security (this document
exists -- to my knowledge -- only in German):\\
\url{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
}
published his prognosis of an almost linear factorization progression,
if all influencing factors are included: Each year the module length feasible for
factorization increases by 20 bit on average. So his forecast was below the
more optimistic estimations of GISA and NIST.

@@

\subsection{Status regarding factorization of concrete large numbers}
\label{nt:NoteFactorization}

@@

\noindent Below the last records are explained in more detail.
The two methods, GNFS and SNFS, used to do so are shortly illustrated at the 
following web pages:
\index{General Number Field Sieve (GNFS)}
\index{Special Number Field Sieve (SNFS)}
\vspace{-10pt}
\begin{itemize}
\item[]
      \url{http://en.wikipedia.org/wiki/Special_number_field_sieve} \\
      \url{http://en.wikipedia.org/wiki/General_number_field_sieve}
\end{itemize}
%\vspace{-10pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
\url{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html}
}
factorized a 158-digit decimal number into its both prime factors (these are
build with 73 and 86 decimal digits) using the GNFS method (General Number
Field Sieve)\index{General Number Field Sieve (GNFS)}.

@@

The task of the researchers from Bonn was not initiated by a challenge, but
they wanted to find the last prime factors of the integer $2^{953} - 1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \url{http://www.cerias.purdue.edu/homes/ssw/cun/}}).

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g.\ using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorization of a Number}. \\
CrypTool can factorize in a reasonable time numbers no longer than 250 bit.
Numbers bigger than 1024 bits are currently not accepted by CrypTool.
}.
The next three prime factors were found by P.~Zimmerman\footnote{%
\url{http://www.loria.fr/~zimmerma/ecmnet}}, 
T.~Grandlund\footnote{\url{http://www.swox.se/gmp/}}
and R. Harley during the years 1999 and 2000 using the elliptic curve
factorization method.

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \url{http://www.loria.fr/~zimmerma/records/gnfs158}\\
          \url{http://www.crypto-world.com/FactorRecords.html}\\
          \url{http://www.crypto-world.com/announcements/c158.txt}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
          \url{http://www.loria.fr/~zimmerma/records/rsa160} \\
          \url{http://www.loria.fr/~zimmerma/records/factor.html} \\
          \url{http://www.crypto-world.com/FactorWorld.html}
} 
factorized a 160-digit number into its both prime factors (these are build
with each 80 decimal digits) using the GNFS method (General Number Field
Sieve)\index{General Number Field Sieve (GNFS)}.

The  computations for the factorization of RSA-160 also took place at the 
German Information Security Agency (GISA) in Bonn\footnote{%
Every year the GISA\index{GISA} creates a paper to describe which 
crypto algorithms are feasible to generate digital signatures according
to the German signature law -- under participation of experts from
economy and science. To review signature methods based on the
factorization problem the GISA also co-operates with researchers from the
University of Bonn.
Further information about crypto algorithms can be found on the web page of GISA:
   \url{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
}.

@@

The task of the researchers was not initiated by a challenge, but
they wanted to find the last prime factors of the Mersenne number $2^{1039}+1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \url{http://www.cerias.purdue.edu/homes/ssw/cun/}\\
Cunningham table: \url{http://homes.cerias.purdue.edu/~ssw/cun/pmain1206}\\
The numbers in the Cunningham table have the following syntax:\\
``(2,n)-'' means $2^{n}-1$;~~~
``(2,n)+'' means $2^{n}+1$.\\
To describe the magnitude one writes $p<n>$ or $c<n>$: 
``n'' is the number of decimal digits and ``p'' and ``c'' tell,
whether the number is prime or composite.\\
$2^{1039}-1 = p7 * c307 = p7 * p80 * p227$ \\
It is explained more precisely at the CUN page:\\
``2651+ means $2^{651} + 1$ and the size (c209 means 209 decimal digits)
of the
number which was factored.  Then come the new factor(s), the discoverer and
the method used.  Recently, only the multiple polynomial quadratic sieve
(ppmpqs), the elliptic curve method (ecm) and the number field sieve (nfs)
have been used.  `hmpqs' stands for hypercube multiple polynomial quadratic
sieve.  Under `new factors', `p90' means a 90-digit prime and `c201' is a
201-digit composite number.''.
}).\\

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]]   \url{http://www.loria.fr/~zimmerma/records/21039-}\\
          \url{http://www.crypto-world.com/announcements/m1039.txt}\\
          \url{http://www.crypto-world.com/FactorAnnouncements.html}\\
          \url{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?detailjahr=2007&detail=160}
\end{itemize}
\end{minipage}

@@

In his paper ``Circuits for integer factorization: a proposal'' 
(\url{http://cr.yp.to/djb.html), published November 2001,
D.~J.\ Bernstein \cite{nt:Bernstein2001} addresses the problem of
factorizing large integers. Therefore, his results are of relevance from a
RSA point of view.  As a main result Bernstein claims that the
implementation of the General Number Field Sieve algorithm (GNFS)
 \index{General Number Field Sieve (GNFS)} can be improved to factor, with
the same effort as before, integers with three times more digits.

@@

RSA Security\footnote{\url{http://www.rsasecurity.com/}} concludes in its
analysis of the Bernstein paper \cite{nt:RSA Security 2002} from April, 8 2002
also -- as expected -- that RSA is still not compromised.

@@

\vskip +12pt
\noindent\begin{minipage}{\textwidth}
Links:
\vspace{-10pt}
\begin{itemize}
  \item[] \url{http://cr.yp.to/djb.html}\\
          \url{http://www.counterpane.com/crypto-gram-0203.html\#6} \\
          \url{http://www.math.uic.edu}
\end{itemize}
\end{minipage}

@@

\noindent Current research results on this topic can be found at:
\vspace{-10pt}
\begin{itemize}
  \item[] \url{http://www.mersenne.org/} \\
          \url{http://fatphil.org/maths/AKS/} Original paper in English\\
          \url{http://ls2-www.cs.uni-dortmund.de/lehre/winter200203/kt/material/primes.ps} \\
	  \hspace*{2em}Good explanation in German by Thomas Hofmeister.
\end{itemize}
\vskip +10 pt

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{Chapter_ElementaryNT_12}{}
\section[The RSA procedure with actual numbers]{The RSA procedure with actual numbers\footnotemark}
\footnotetext{%
    \index{Sage}%
    \index{Nguyen, Minh Van }%
    Additional material: Minh Van Nguyen, ``Number Theory and the RSA
    Public Key Cryptosystem'', 2009. An introductory tutorial on using
    Sage to study elementary number theory and public key
    cryptography. A didactically very clear article about some basic
    number theory and Sage usage. 
    \url{http://nguyenminh2.googlepages.com/sage_numtheory-rsa.pdf}.
}
\label{rsaconcrete}\index{RSA}
Having described above \hyperlink{RSA}{how the RSA procedure works}, we will now work through the steps using actual, but small, numbers.

@@

The task is taken from \cite[Exercise 4.6]{nt:Stinson1995}\index{Stinson 1995}:
The pure solution has been published by Prof. Stinson at 
\url{http://www.cacr.math.uwaterloo.ca/~dstinson/solns.html}\footnote{%
or
\url{http://bibd.unl/~stinson/solns.html}.
}

@@

The file ``primroots.dat'' is a database of all primitive roots of all primes
between 1 and 100,000 inclusive. It is a very large file (about 1 GB
uncompressed, and 285 MB compressed with bzip2). You can find the file
at
   \href{http://sage.math.washington.edu/home/mvngu/doc/primitive-roots/primroots.dat.bz2}  %%% WARUM DOPPELT ????? be_be
   {\url{http://sage.math.washington.edu/home/mvngu/doc/primitive-roots/primroots.dat.bz2}}.

@@

\begin{enumerate}
   \item \hypertarget{knott}{}  \index{Knott, Ron}
         Ron Knott's Fibonacci \index{Fibonacci} page, \\
         Here, everything revolves around Fibonacci numbers.\\
          \url{http://www.mcs.surrey.ac.uk/personal/R.Knott/Fibonacci/fib.html}

   \item CrypTool, \index{CrypTool} \\
         Open source e-learning software to illustrate cryptography and
         cryptanalysis\\
          \url{http://www.cryptool.de}, \\
          \url{http://www.cryptool.org},\\ 
          \url{http://www.cryptool.com}

   \item Mathematica, \index{Mathematica} \\
         Commercial mathematics package \\
         \url{http://www.wolfram.com}

   \item LiDIA, \index{LiDIA} \\
         Extensive library containing number-theory functions and the
         LC interpreter. Maintenance stopped. \\
          \url{http://www.informatik.tu-darmstadt.de/TI/LiDIA}

   \item BC, \index{BC} \\
         Interpreter with number-theory functions \\
         \url{http://www.gnu.org/software/bc/bc.html}

   \item Pari-GP, \index{Pari-GP} \\
         Excellent, fast, free interpreter with number theoretical functions.\\
         \url{http://pari.math.u-bordeaux.fr/} \\
         \url{http://en.wikipedia.org/wiki/PARI/GP} \\
         Resources for PARI/GP at Karim Belabas's website:\\
         \url{http://www.math.u-bordeaux.fr/~belabas/pari/}

   \item \index{Munchenbach@M\"unchenbach, Carsten}
         Only after I had completed this article, did I come across the
         website of Mr.\ M\"unchenbach, which interactively and didactically
         uses elementary number theory to provide a sophisticated
         description of the fundamental mathematical thought processes. 
         It was created for a teaching project in the 11th grade of the
         technical grammar school (unfortunately only available in German): \\
         \url{http://www.hydrargyrum.de/kryptographie}

   \item Web site of Mr.~Wagner, who is responsible for the development
	 of the curriculum of computer science in one of the German federal 
         states (L\"ander). Here you can get hold of a collection of texts
	 and (Java-)\discretionary{}{}{}programs (available only in German):\\
         \url{http://www.saar.de/~awa/kryptolo.htm}

   \item GISA, \index{GISA} \\
         German Information Security Agency \\
         \url{http://www.bsi.bund.de}

   \item Factorization records and challenges\index{Factorization!factoring records},\\
         \url{http://www.crypto-world.com/} \\
         \url{http://www.crypto-world.com/FactorWorld.html}, Webseite von Scott Contini \\
         \url{http://www.loria.fr/~zimmerma/records/factor.html} \\

         \url{http://www.tutorgig.com/ed/RSA\_number} \\

         \url{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html} \\
         \url{http://www.ercim.org/publication/Ercim\_News/enw49/franke.html} \\ %2002-01
         \url{http://www.loria.fr/~zimmerma/records/rsa160} \\

         \url{http://www.rsa.com/rsalabs/node.asp?id=2092}

   \item The Cunningham Project, \index{Cunningham project}\\ 
         \url{http://www.cerias.purdue.edu/homes/ssw/cun/}

   \item Sage, \index{Sage} \\
         Excellent, open source computer algebra system with Python as script language,
	 used to build the code samples in this chapter.
	 See the introduction in chapter~\ref{s:appendix-using-sage}.\\
         \url{http://www.sagemath.org/} \\
         \url{http://en.wikipedia.org/wiki/Sage_%28mathematics_software%29}
